DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-04-15 paper/comparison.tex
Updates to paper

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/compare_benchmark}
  \caption{Average control loop execution time in the presence of an
    increasing number of memory intensive applications on idle CPU cores.}
  \label{fig:sys_bench}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/compare_l2missrate}
  \caption{L2 miss rates in the presence of an increasing number of 
			memory intensive applications on idle CPU cores.}
  \label{fig:sys_l2miss}
\end{figure}

Finally, we compare the effect of co-scheduling memory bandwidth
intensive synthetic benchmarks on the DNN control loop timing. 
Figure~\ref{fig:sys_bench} shows the results. As discussed in 
Section~\ref{sec:eval-memhog}, we observed dramatic execution time
increases, up to 11.6X, in Raspberry Pi 3 as we increased the number of
co-scheduled tasks. We also observe increased control loop execution
timing in the Intel Up and Jetson TX2, but the degree of the increase is 
not as dramatic as the Pi 3. Compared to their respective solo timings 
(i.e., the model runs on a single core in isolation), Intel UP suffers up to
2.3X execution time increase; TX2 (CPU) and TX2 (GPU) suffer up to
2.2X and 2.1X increases, respectively. This is somewhat surprising
because the Raspberry Pi 3's cores are in-order architecture based while
the cores in the Intel Up and NVIDIA TX2 are out-of-order architecture
based, and that the memory intensive tasks on out-of-order cores can
generate more memory traffic. We believe that this is because the
memory subsystems in the Intel UP and TX2 platforms provide higher
performance than the memory subsystem of the Pi 3 as suggested from
the measured memory bandwidth results in Table~\ref{tbl:platforms}
('Measured Bandwidth').

Another interesting observation is that the TX2 (GPU) also suffers
considerable execution time increase (2.1X) despite the fact that the
co-scheduled synthetic tasks do not utilize the GPU. In other words,
the DNN model has dedicated access to the GPU. This is, however, a
known characteristic of integrated CPU-GPU architecture based
platforms where both CPU and GPU share the same memory
subsystem~\cite{Ali2017}. As a result, the TX2 (GPU) fails to meet the
10ms deadline for 100 Hz control that would have been feasible if
there was no contention between the CPU cores and the GPU.

Figure~\ref{fig:sys_l2miss} shows the L2 miss rates of DNN control
loop. The key takeaway of this result is that the L2 miss rate is a
poor indicator to the execution time of the control loop. As shown
already, the Raspberry Pi 3's increased L2 misses do not fully explain
increases in processing times. In case of the Intel UP, as the number of
interfering Bandwidth benchmark instances increases, the L2 miss rates
drop slightly. In the case of the TX2 (CPU), from solo to one co-runner,
L2 miss rates almost double, but then the execution time
increase is a little less than 10\%. However, when there are
co-runners, even though the L2 miss rates are largely unchanged, the
DNN processing time increases significantly. The TX2 (GPU) also shows
similar behavior. All of these results essentially point to the fact
that DNN inferencing workload is largely memory performance sensitive
but not L2 cache space sensitive.

>>>>

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/compare_benchmark_read}
  \caption{Average control loop execution time in the presence of an
    increasing number of memory intensive read applications on idle CPU cores.}
  \label{fig:sys_bench_read}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/compare_benchmark}
  \caption{Average control loop execution time in the presence of an
    increasing number of memory intensive write applications on idle CPU cores.}
  \label{fig:sys_bench}
\end{figure}

Finally, we compare the effect of co-scheduling memory bandwidth
intensive synthetic benchmarks on the DNN control loop timing. 
Figure~\ref{fig:sys_bench_read} and ~\ref{fig:sys_bench} shows 
the results for read and write co-runners, respectively. For read co-runners,
the performance of all platforms decreased as additional BwRead processes
were introduced, but none of them displayed any significant or unexpected
execution time increases. The Pi 3 suffered an increase of up to 1.6X; the 
Intel UP increased up to 1.3X; the TX2 (CPU) and TX2(GPU) saw increases 
up to1.6X and 2.2X, respectively. For write co-runners, as discussed in 
Section~\ref{sec:eval-memhog}, we observed dramatic execution time
increases, up to 11.6X, in the Raspberry Pi 3 as we increased the number of
co-scheduled tasks. We also observe increased control loop execution
timing in the Intel Up and Jetson TX2, but the degree of the increase is 
not as dramatic as the Pi 3. Compared to their respective solo timings 
(i.e., the model runs on a single core in isolation), Intel UP suffers up to
2.3X execution time increase; TX2 (CPU) and TX2 (GPU) both suffer up to
2.3X increases. This is somewhat surprising
because the Raspberry Pi 3's cores are in-order architecture based while
the cores in the Intel Up and NVIDIA TX2 are out-of-order architecture
based, and that the memory intensive tasks on out-of-order cores can
generate more memory traffic. We believe that this is because the
memory subsystems in the Intel UP and TX2 platforms provide higher
performance than the memory subsystem of the Pi 3 as suggested from
the measured memory bandwidth results in Table~\ref{tbl:platforms}
('Measured Bandwidth').

Another interesting observation is that the TX2 (GPU) also suffers
considerable execution time increase (2.3X) despite the fact that the
co-scheduled synthetic tasks do not utilize the GPU. In other words,
the DNN model has dedicated access to the GPU. This is, however, a
known characteristic of integrated CPU-GPU architecture based
platforms where both CPU and GPU share the same memory
subsystem~\cite{Ali2017}. As a result, the TX2 (GPU) fails to meet the
10ms deadline for 100 Hz control that would have been feasible if
there was no contention between the CPU cores and the GPU.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/compare_readl2}
  \caption{L2 miss rates in the presence of an increasing number of 
			memory intensive read applications on idle CPU cores.}
  \label{fig:sys_readl2miss}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/compare_l2missrate}
  \caption{L2 miss rates in the presence of an increasing number of 
			memory intensive write applications on idle CPU cores.}
  \label{fig:sys_l2miss}
\end{figure}

Figures~\ref{fig:sys_readl2miss} and ~\ref{fig:sys_l2miss} show
the L2 miss rates of DNN control loops when read and write co-runners
are introduced, respectively. The key takeaway of this result is that the 
L2 miss rate is a
poor indicator to the execution time of the control loop. As shown
already, the Raspberry Pi 3's increased L2 misses do not fully explain
increases in processing times. In case of the Intel UP, as the number of
interfering Bandwidth benchmark instances increases, the L2 miss rates
drop slightly. In the case of the TX2 (CPU), from solo to one co-runner,
L2 miss rates almost double, but then the execution time
increase is a little less than 10\%. However, when there are
co-runners, even though the L2 miss rates are largely unchanged, the
CNN processing time increases significantly. The TX2 (GPU) also shows
similar behavior. All of these results essentially point to the fact
that DNN inferencing workload is largely memory performance sensitive
but not L2 cache space sensitive.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/compare_readbus}
  \caption{Bus accesses in the presence of an increasing number of 
			memory intensive read applications on idle CPU cores.}
  \label{fig:sys_readbus}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/compare_writebus}
  \caption{Bus accesses in the presence of an increasing number of 
			memory intensive write applications on idle CPU cores.}
  \label{fig:sys_writebus}
\end{figure}

\color{red}FIXME: I believe the bus access results for the Intel UP are wrong. 
I couldn't get the IDI off-core response event to work, so I used the bus-cycles 
perf event, which I think is incorrect.

\color{black}Figures~\ref{fig:sys_readbus} and ~\ref{fig:sys_writebus} show the
normalized bus accesses for each of the platforms as read and write
co-runners are introduced, respectively. Unlike the L2 miss rates of the
model, the total number of bus accesses more closely align with the 
execution time increases that are seen in each of the platforms. For the Pi 3, both 
the execution time and bus accesses increase at exponential rates. In the TX2(GPU), 
the bus accesses increase in a similar linear fashion to the execution time. The main outlier
is the TX2(CPU) which has bus accesses increase at a faster rate similar to that of the Pi 3.
However, the bus accesses seem to increase linearly, which is also seen in the
execution time. As a result, we find that the total number of bus accesses have a higher 
correlation to the model execution time than L2 miss rates.

### 2018-04-15 paper/conclusion.tex
Updates to paper

\section{Conclusion}\label{sec:conclusion}
We presented DeepPicar, a low cost autonomous car platform that is
inexpensive to build, but is based on state-of-the-art AI technology:
End-to-end deep learning based real-time control.
Specifically, DeepPicar uses a deep convolutional neural network to
predict steering angles of the car directly from camera input data
in real-time. Importantly, DeepPicar's neural network architecture is
identical to that of NVIDIA's real self-driving car.

@@

As future work, we plan to apply shared DRAM resource management
techniques~\cite{Yun2013,yun2014rtas} on the DeepPicar platform and
evaluate their impact on the real-time performance of the system. We
also plan to improve the prediction accuracy by feeding more data and
upgrading the RC car hardware platform to enable more precise steering
angle control.

>>>>

\section{Conclusion}\label{sec:conclusion}
We presented DeepPicar, a low cost autonomous car platform that is
inexpensive to build, but is based on state-of-the-art AI technology:
End-to-end deep learning based real-time control.
Specifically, DeepPicar uses a deep convolutional neural network to
predict steering angles of the car directly from camera input data
in real-time. Importantly, DeepPicar's neural network architecture is
almost identical to that of NVIDIA's real self-driving car.

@@

As future work, 
%% we plan to apply shared DRAM resource management
%% techniques~\cite{Yun2013,yun2014rtas} on the DeepPicar platform and
%%evaluate their impact on the real-time performance of the system. 
we plan to improve the prediction accuracy by feeding more data and
upgrading the RC car hardware platform to enable more precise steering
angle control.

### 2018-04-15 paper/evaluation-cachepart.tex
Updates to paper

In order to validate the effectiveness of cache partitioning, we patch
the kernel with PALLOC~\cite{yun2014rtas}, which is a kernel-level
color-aware page allocator, supporting cache partitioning. 
Figure~\ref{fig:cache-mapping} shows the physical address mapping of
Raspberry Pi3's BCM2837 processor. We use the bit 12, 13, 14 for
coloring, which results in 8 page colors.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/palloc_multicore}
  \caption{Average control loop execution time vs.  % of L2 cache
    L2 cache partition size.}
  \label{fig:palloc_multicore}
\end{figure}

In the first experiment, we investigate the cache space sensitivity of
the DeepPicar's DNN-based control loop. Using PALLOC, we create 4
different cgroups which are configured to use 2, 4,
6, and 8 colors (25\%, 50\%, 75\%, and 100\% of the L2 cache
space, respectively). We then execute the control loop on each cache
partition (cgroup) and measure the average processing time.

If performance improves as a result of partitioning the shared cache 
then we know that the DNN, to some extent, relies on shared cache 
performance in addition to shared DRAM performance. On the other 
hand, if there is no improvement in performance, then it can be 
observed that the shared memory is more critical to the performance 
of the DNN.
%% We run the DNN model with different amounts of L2 cache being
%% available based on the number of colors being used. For example, if two 
%% colors are used, then 25\% of the L2 cache is available, whereas if all 
%% 8 colors are used, 100\% of the L2 cache is available. Particularly, we 
%% test the performance of the model when it utilizes a single core and
%% all four of the available cores. 
The results can be seen in 
\ref{fig:palloc_multicore}. As expected, the amount of L2 cache 
available to the model does not result in any significant changes in 
performance. 
%% This can also be seen by the L2 miss rates of the model 
%% seen in \ref{fig:palloc_multicore_l2missrate}. As more cache space 
%% becomes avaiable, the percentage of L2 misses decreases, which doesn't 
%% affect the real-time performance of the DNN. As such, we find that a 
%% single model is not sensitive to the shared L2 cache.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/palloc_multimodel}
  \caption{Timing impact of co-scheduling multiple DNNs when cache 
partitioning is enabled.}
  \label{fig:palloc_multimodel}
\end{figure}

We also co-schedule multiple models to see how cache partitioning affects
interference between them. Each model uses an equal number of colors, and
are thus given equal amounts of L2 cache space, and there is no overlap
in the colors used, so no L2 cache is shared between models. As can be 
seen in Figure \ref{fig:palloc_multimodel}, the performance of the models 
remains the same as when no cache partitioning was employed. Based of 
these results, we find that contention for the L2 cache is not the main
source of interference between multiple DNN models.

@@

Finally, we run a single model alongside memory intensive co-runners 
again, but ensure that each task is given equal amounts of L2 cache space.
For this, we assign two colors to each task, regardless of the number of
tasks being scheduled. As a result, the entire cache is not utilized in
experiments with less than four tasks. The results can be seen in Figure
\ref{fig:palloc_bandwidth_exectime}. When the model runs by itself, performance
is slightly worse due to less cache space being available (this was also seen in 
Figure \ref{fig:palloc_multicore}). With one or two co-runners, the DNN
performance doesnt't change relative to when the cache wasn't partitioned.
When three co-runners are present, DNN performance becomes slightly
better, with a improvement of \textasciitilde6.5\%. However, as the 
performance improvement is minimal, we do not believe that cache 
partitioning is effective for protecting the model from interference from
other memory intensive processes.
%% At the same time, though, the L2 miss rates of the model don't correlate 
%% to the timing increases seen, as is shown in 
%% \ref{fig:palloc_bandwidth_l2missrate}. Instead, we find that the number
%% of bus accesses increases when cache partitions are used, and they mirror
%% the timing increases of the model. This can be seen in Figure 
%% \ref{fig:palloc_bandwidth_bus}. These additional bus accesses are
%% most likely caused by the L2 prefetcher, as it can generate additional bus
%% accesses that aren't counted as L2 misses. As a result, we find that cache
%% partitioning is detrimental to the performance of the DNN so long as the 
%% 0L2 prefetcher is enabled (we have not yet found a way to disable it on
%% the Pi 3). Furthermore, no cache sensitivity was displayed by the model,
%% thus showing that the performance of the L2 cache is not vital for DNN 
%% performance when co-runners are introduced.

By partitioning the shared L2 cache, we find that no noticable 
improvements are gained and that performance remains consistent, or 
slightly decreases, depending on the circumstances. In all experiements,
the DNN shows no sensitivity to the shared L2 cache. As a result, we 
conclude that cache partitioning is not an effective isolation mechanism, 
and that the performance of the shared DRAM controller is of greater 
importance to the real-time efficiacy of the DeepPicar.

>>>>

In order to validate the effectiveness of cache partitioning, we patch
the kernel with PALLOC~\cite{yun2014rtas}, which is a kernel-level
color-aware page allocator, supporting cache partitioning. 
Figure~\ref{fig:cache-mapping} shows the physical address mapping of
Raspberry Pi3's BCM2837 processor. We use the bits 13 and 14 for
coloring, which results in 4 page colors.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/palloc_multicore}
  \caption{Cache space sensitivity of the control loop exectuion time.}
  \label{fig:palloc_multicore}
\end{figure}

In the first experiment, we investigate the cache space sensitivity of
the DeepPicar's DNN-based control loop. Using PALLOC, we create 4
different cgroups which are configured to use 1, 2,
3, and 4 colors (25\%, 50\%, 75\%, and 100\% of the L2 cache
space, respectively). We then execute the control loop on each cache
partition (cgroup) and measure the average processing time.

If performance improves as a result of partitioning the shared cache 
then we know that the DNN, to some extent, relies on the shared cache 
in addition to the shared DRAM. On the other 
hand, if there is no improvement in performance, then it can be 
observed that the shared memory is more critical for the DNN.
%% We run the DNN model with different amounts of L2 cache being
%% available based on the number of colors being used. For example, if two 
%% colors are used, then 25\% of the L2 cache is available, whereas if all 
%% 8 colors are used, 100\% of the L2 cache is available. Particularly, we 
%% test the performance of the model when it utilizes a single core and
%% all four of the available cores. 
The results can be seen in 
\ref{fig:palloc_multicore}. As expected, the amount of L2 cache 
available to the model does not result in any significant changes in 
performance. 
%% This can also be seen by the L2 miss rates of the model 
%% seen in \ref{fig:palloc_multicore_l2missrate}. As more cache space 
%% becomes avaiable, the percentage of L2 misses decreases, which doesn't 
%% affect the real-time performance of the DNN. As such, we find that a 
%% single model is not sensitive to the shared L2 cache.

%% \begin{figure}[h]
%%  \centering
%%   \includegraphics[width=.45\textwidth]{figs/palloc_multimodel}
%%   \caption{Timing impact of co-scheduling multiple DNNs when cache 
%% partitioning is enabled.}
%%   \label{fig:palloc_multimodel}
%% \end{figure}

%% We also co-schedule multiple models to see how cache partitioning affects
%% interference between them. Each model uses an equal number of colors, and
%% are thus given equal amounts of L2 cache space, and there is no overlap
%% in the colors used, so no L2 cache is shared between models. As can be 
%% seen in Figure \ref{fig:palloc_multimodel}, the performance of the models 
%% remains the same as when no cache partitioning was employed. Based of 
%% these results, we find that contention for the L2 cache is not the main
%% source of interference between multiple DNN models.

@@

We also run a single model alongside three memory intensive read co-runners 
again, but ensure that each task is given equal amounts of L2 cache space.
For this, we assign one color to each task. The results can be seen in Figure
\ref{fig:palloc_bandwidth_exectime}. When cache partitioning is enabled, 
DNN performance becomes slightly better, with a improvement of 
\textasciitilde6.5\%. However, as the performance improvement is minimal, 
we do not believe that cache partitioning is effective for protecting the model 
from interference from other memory intensive processes.
%% At the same time, though, the L2 miss rates of the model don't correlate 
%% to the timing increases seen, as is shown in 
%% \ref{fig:palloc_bandwidth_l2missrate}. Instead, we find that the number
%% of bus accesses increases when cache partitions are used, and they mirror
%% the timing increases of the model. This can be seen in Figure 
%% \ref{fig:palloc_bandwidth_bus}. These additional bus accesses are
%% most likely caused by the L2 prefetcher, as it can generate additional bus
%% accesses that aren't counted as L2 misses. As a result, we find that cache
%% partitioning is detrimental to the performance of the DNN so long as the 
%% 0L2 prefetcher is enabled (we have not yet found a way to disable it on
%% the Pi 3). Furthermore, no cache sensitivity was displayed by the model,
%% thus showing that the performance of the L2 cache is not vital for DNN 
%% performance when co-runners are introduced.

By partitioning the shared L2 cache, we find that no noticable 
improvements are gained and that performance remains consistent. 
In all experiements, the model shows no sensitivity to the shared L2 
cache. As a result, we conclude that cache partitioning is not an 
effective isolation mechanism, and that the performance of the shared 
DRAM controller is of greater importance to the real-time efficiacy of 
the DeepPicar.

### 2018-04-15 paper/evaluation-drampart.tex
Updates to paper

To determine the how sensitive the model is to the shared DRAM 
memory, we use MemGuard, a memory bandwidth reservation system that 
can limit the amount of bandwidth each core receives in a period. For 
all tests run using MemGuard, we use a period of one second.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/memguard_multicore}
  \caption{Average control loop execution time vs.
    memory bandwidth reservation size. }
  \label{fig:memguard_multicore}
\end{figure}

We first run a single model on one core, Core 0, while varying the 
core's bandwidth reservation size, from 400 MB/s down to 100 MB/s. 
The results can be seen in Figure \ref{fig:memguard_multicore}. When less 
memory bandwidth is available to the core the model's performance 
noticably decreases, and it performs the best when no memory bandwidth 
throttling is implemented. Compared to a reservation size of 100 
MB/s, disabling memory bandwidth throttling results in 
\textasciitilde3x performance improvement. If the model was memory 
insensitive, then the amount of bandwidth avaiable to it would have 
little effect on its performance. However, since that is not the 
case, we conclude that, to a notable extent, the neural network model 
is dependent on shared DRAM memory performance.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/memguard_multimodel}
  \caption{Timing impact of co-scheduling multiple DNNs when memory
bandwidth throttling is enabled. }
  \label{fig:memguard_multimodel}
\end{figure}

We also test the effects of memory bandwidth throttling in the case 
of multiple models running concurrently on the Pi 3 by rerunning the 
4Nx1C experiment. We use the same memory bandwidth reservation sizes 
from the previous experiment. The results can be seen in Figure
\ref{fig:memguard_multimodel}. Once again, the performance of the 
models are affected by the amount of memory bandwidth that is 
available to the Pi 3's cores during each period, meaning that they 
are all memory dependent.

@@

Finally, we test the performance of the model in the presence of 
memory intensive read co-runners in order to determine the effects of 
limiting the memory bandwidth of the co-runners. For each of the 
experiments, we give the model's core, Core 0, a constant 1000 MB/s 
memory bandwidth reservation, and then vary the amount of bandwidth 
the co-runners' cores receive following the same scheme used in the 
previous experiments. The results can be see in Figure
\ref{fig:memguard_bandwidth}. When less memory bandwidth is given to 
the read co-runners, the performance of the model does improve. In 
the case of one co-runner, the performance of the model remains 
constant even as the reservation sizes are decreased, however, it 
still improves compared to when throttling wasn't enabled. With two 
and three co-runners, performance continually improves as less 
bandwidth is given to the benchmarks. This is especially true with 
three co-runners as the improvement is linear, which indicates that 
there is a direct correlation between the co-runner bandwidth size 
and model performance. In all cases, model performance improved, to 
some extent, through the use of memory bandwidth throttling on the 
co-runners. Since this would not be the case if the model was memory 
insensitive, we find that the model is memory dependent in the 
presence of co-runners.

Based on all of the memory bandwidth throttling experiments 
performed, it is evident that the model is dependent on the shared 
DRAM memory, to a noticable extent. When a single or multiple models 
were run, their performance(s) decreased as less memory bandwidth was 
made available to them. Furthermore, when the bandwidths of memory 
intensive read co-runners were limited, the performance of the model 
increased. As such, we conclude that the model is dependent on the 
performance of the shared DRAM memory.

>>>>

To determine the how sensitive the model is to the shared DRAM 
memory, we use MemGuard\cite{yun2014rtas}, a memory bandwidth reservation system that 
can limit the amount of bandwidth each core receives in a period. For 
all tests run using MemGuard, we use a period of one second.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/memguard_multicore}
  \caption{ Memory bandwidth sensitivity of the control loop 
execution time. }
  \label{fig:memguard_multicore}
\end{figure}

We first run a single model on one core, Core 0, while varying the 
core's bandwidth reservation size, from 500 MB/s down to 100 MB/s. 
The results can be seen in Figure \ref{fig:memguard_multicore}. When less 
memory bandwidth is available to the core the model's performance 
noticably decreases.
%% , and it performs the best when no memory bandwidth 
%% throttling is implemented. 
Compared to a reservation size of 100 
MB/s, disabling memory bandwidth throttling results in 
\textasciitilde3x performance improvement. If the model was memory 
insensitive, then the amount of bandwidth available to it would have 
little effect on its performance. However, since that is not the 
case, we conclude that, to a notable extent, the neural network model 
is dependent on shared DRAM memory performance.

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.45\textwidth]{figs/memguard_multimodel}
%%   \caption{Timing impact of co-scheduling multiple DNNs when memory
%% bandwidth throttling is enabled. }
%%   \label{fig:memguard_multimodel}
%% \end{figure}

%% We also test the effects of memory bandwidth throttling in the case 
%% of multiple models running concurrently on the Pi 3 by rerunning the 
%% 4Nx1C experiment. We use the same memory bandwidth reservation sizes 
%% from the previous experiment. The results can be seen in Figure
%% \ref{fig:memguard_multimodel}. Once again, the performance of the 
%% models are affected by the amount of memory bandwidth that is 
%% available to the Pi 3's cores during each period, meaning that they 
%% are all memory dependent.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/memguard_bwwrite}
  \caption{Timing impact of co-scheduling memory intensive write
co-runners when memory bandwidth throttling is enabled. }
  \label{fig:memguard_bwwrite}
\end{figure}

We also test the performance of the model in the presence of three
memory intensive read and write benchmarks to determine the effects of 
limiting the memory bandwidth of the co-runners. For each of the 
experiments, we give the model's core, Core 0, a constant 1000 MB/s 
memory bandwidth reservation size, and then vary the amount of bandwidth 
the co-runners' cores receive. For read co-runners, we give them an initial
reservation size of 500 MB/s, and decrease that to a minimum of 100 MB/s.
In the case of write co-runners, MemGuard accounts for L2 refills but not
write-backs, so each benchmark would effectively have twice the bandwidth
of their reservation size. Due to this, we give write co-runners an initial 
reservation size of 250 MB/s and decrease it to a minimum of 50 MB/s.
The results can be see in Figures \ref{fig:memguard_bandwidth} and
\ref{fig:memguard_bwwrite}. When less memory bandwidth is given to 
the both the  read and write co-runners, the performance of the model 
improves. 
%% In the case of one co-runner, the performance of the model remains 
%% constant even as the reservation sizes are decreased, however, it 
%% still improves compared to when throttling wasn't enabled. 
In both cases, when the co-runners were given a minimal reservation size,
the performance of the model was much closer to its solo execution time.
This is especially noteworthy when write co-runners were present, as the 
model didn't suffer the same 11.6X slowdown that was seen without
bandwidth throttling enabled. 
%% When memory intensive Bandwidth benchmarks were present, model performance
%% improved, to a notable extent, through the use of memory bandwidth 
%% throttling on the co-runners. 
Since this would not be the case if the 
model was memory insensitive, we find that the model is memory 
dependent in the presence of co-runners.

Based on all of the memory bandwidth throttling experiments 
performed, it is evident that the shared DRAM memory is important for
the DNN model. When a single model was run, its 
performance decreased as less memory bandwidth was 
made available to it. Furthermore, when the bandwidths of memory 
intensive read and write co-runners were limited, the performance of 
the model improved. As such, we conclude that the model is dependent on the 
shared DRAM memory.

### 2018-04-15 paper/evaluation.tex
Updates to paper

\begin{figure}[t]
  \centering
  \includegraphics[width=.5\textwidth]{figs/Fig7_new}
  \caption{DeepPicar's control loop processing times over 1000 input image frames.}
  \label{fig:control-loop-timing}
\end{figure}

\begin{table}[t]
  \centering
  \begin{tabular} {| c | r | r | r | r |}
    \hline
    \textbf{Operation} & \textbf{Mean} & \textbf{Max} &   \textbf{99pct.} & \textbf{Stdev.} \\ \hline
    Image capture        & 1.61  &  1.81 &  1.75  & 0.05 \\ \hline
    Image pre-processing & 2.77  &  2.90 &  2.87  & 0.04 \\ \hline
    DNN inferencing      & 18.49 & 19.30 & 18.99  & 0.20 \\ \hline
    Total Time           & 22.87 & 23.74 & 23.38  & 0.20 \\ \hline
  \end{tabular}
  \caption{Control loop timing breakdown.}
  \label{tbl:control-loop-breakdown}
\end{table}

@@

@@

Second, we find that scalability of Tensorflow's DNN 
implementation is limited. We find that using all four cores can double 
performance of the DNN, but compared to using three cores, performance
improvement is relatively minimal.

@@

Fifth, we find that partitioning the shared L2 cache provides 
practically no benefits to the real-time performance of the DNN. To the 
contrary, performance is constant regardless of the amount of cache 
space used. Furthermore, we find that performance remains consistent 
when multiple models are run simultaneously, and when memory intensive
read co-runners are introduced. As such, we conclude that 
the shared L2 cache is not essential to the DNN's real-time performance.

Lastly, we find that memory bandwidth throttling results in significant
changes to model performance. When less bandwidth is given to models,
whether they're running alone or concurrently, we find that their performance
decreases noticably. When memory intensive co-runners are
present, we find that limiting the available bandwidth of the co-runners
actually results in improved model performance. As such, we conclude
that the shared DRAM memory is essential to the DNN's real-time 
performance.

>>>>

\begin{figure}[h]
  \centering
  \includegraphics[width=.5\textwidth]{figs/Fig7_new}
  \caption{DeepPicar's control loop processing times over 1000 input image frames.}
  \label{fig:control-loop-timing}
\end{figure}

\begin{table}[h]
  \centering
  \begin{tabular} {| c | r | r | r | r |}
    \hline
    \textbf{Operation} & \textbf{Mean} & \textbf{Max} &   \textbf{99pct.} & \textbf{Stdev.} \\ \hline
    Image capture        & 1.61  &  1.81 &  1.75  & 0.05 \\ \hline
    Image pre-processing & 2.77  &  2.90 &  2.87  & 0.04 \\ \hline
    DNN inferencing      & 18.49 & 19.30 & 18.99  & 0.20 \\ \hline
    Total Time           & 22.87 & 23.74 & 23.38  & 0.20 \\ \hline
  \end{tabular}
  \caption{Control loop timing breakdown.}
  \label{tbl:control-loop-breakdown}
\end{table}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/a53_vs_a7}
  \caption{DNN model performance on different in-order
ARM Cortex cores.}
  \label{fig:a53_vs_a7}
\end{figure}

We do find the performance of the model in the presence of 
write co-runners surprising, as it has been previously shown in experiments
on another in-order ARM core, the Cortex A7, that write interference was not as 
significant\cite{Valsan2016}. We also tested the model, in the presence of write 
co-runners, on the A7 cores of the Odroid XU4 platform to see if  the same model 
behavior would also be seen there. The results can be seen in Figure ~\ref{fig:a53_vs_a7}. 
As expected, when the model is run on A7 cores, the intereference of the write
co-runners is not as impactful on model performance as it can still operate at 
under 10 Hz.

@@

Second, we find that scalability of Tensorflow's DNN 
implementation is limited. We find that using all four cores can double 
performance of the CNN, but compared to using three cores, performance
improvement is relatively minimal.

@@

Fifth, we find that partitioning the shared L2 cache provides 
practically no benefits to the real-time performance of the DNN. To the 
contrary, performance is constant regardless of the amount of cache 
space used. Furthermore, we find that performance remains consistent 
when memory intensive read co-runners are introduced, even though
the model has its own isolated cache space. As such, we 
conclude that the shared L2 cache is not essential to the DNN's real-time 
performance.

Lastly, we find that memory bandwidth throttling results in significant
changes to model performance. When less bandwidth is given to the model,
we find that it performs worse. When memory intensive co-runners are
present, we find that limiting the available bandwidth of the co-runners
results in significantly improved model execution times. As such, we conclude
that the shared DRAM memory is essential to the DNN's real-time 
performance.
